### 2025-02-15 README.Rmd
update dockerfile

---
output: github_document
editor_options: 
  markdown: 
    wrap: 72
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)

```

# Extracting sunbeams from cucumbers: Computational reproducibility is essential for archaeological science

[![.github/workflows/knit-in-docker.yml](https://github.com/benmarwick/web-of-science-archaeology/actions/workflows/knit-in-docker.yml/badge.svg)](https://github.com/benmarwick/web-of-science-archaeology/actions/workflows/knit-in-docker.yml)

This repository contains the data and code for my paper to be submitted
to the Journal of Archaeological Science. Earlier versions were
presented at the University of Cambridge
[(2019)](https://github.com/benmarwick/March-2019-Cambridge-Big-Data-Archaeology)
and New York University
[(2025)](https://github.com/benmarwick/nyu-2025-workshop)

## Contents

The **analysis** directory contains:

-   [:file_folder: paper](/analysis/paper): Quarto source document for
    manuscript. Includes code to reproduce the figures and tables
    generated by the analysis.
-   [:file_folder: data](/analysis/data): Data used in the analysis.
-   [:file_folder: figures](/analysis/figures): Plots and other
    illustrations

## How to download and run locally

This research compendium has been developed using the statistical programming language R. To work with the compendium, you will need installed on your computer the [R software](https://cloud.r-project.org/) itself and optionally [RStudio Desktop](https://rstudio.com/products/rstudio/download/).

You can download the compendium as a zip from from this URL:
[master.zip](/archive/master.zip). After unzipping:

-   open the `.Rproj` file in RStudio\
-   install the packages used in the Quarto document\
-   finally, render the `analysis/paper/paper.qmd` document to produce the `paper.docx`, or run `rmarkdown::render("analysis/paper/paper.qmd")` in the R console

## How to download and run locally in a Docker container

This research compendium includes a [Dockerfile](Dockerfile) that specifies the computational environment used to generate the results presented in the paper. You can recreate this environment on your computer by following these steps:

1. Download the compendium as a zip from from this URL:
[master.zip](/archive/master.zip) and unzip    
2. Install [Docker](https://www.docker.com/get-started/)    
3. Open the compendium in the terminal and run `docker build -t wos .` to build the container on your computer, it will take a few minutes and require an internet connection.
4. Run in your terminal `docker run --rm -it -e ROOT=TRUE -e PASSWORD=rstudio -dp 8787:8787 wos` to run the container
5. Go to http://localhost:8787/ with your browser and enter  rstudio in both fields to start RStudio in the browser, running in the Docker container
6. In RStudio, open the project and the Quarto document, and render it

### Licenses

**Text and figures :**
[CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)

**Code :** See the [LICENSE](LICENSE.md) file

**Data :** [CC-0](http://creativecommons.org/publicdomain/zero/1.0/)
attribution requested in reuse

### Contributions

We welcome contributions from everyone. Before you get started, please
see our [contributor guidelines](CONTRIBUTING.md). Please note that this
project is released with a [Contributor Code of Conduct](CONDUCT.md). By
participating in this project you agree to abide by its terms.