### 2025-02-15 README.Rmd
dockerfile and instructions

---
output: github_document
editor_options: 
  markdown: 
    wrap: 72
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)

```

# Extracting sunbeams from cucumbers: Computational reproducibility is essential for archaeological science

This repository contains the data and code for my paper to be submitted to the Journal of Archaeological Science. Earlier versions were presented at the University of Cambridge [(2019)](https://github.com/benmarwick/March-2019-Cambridge-Big-Data-Archaeology) and New York University [(2025)](https://github.com/benmarwick/nyu-2025-workshop)

## Contents

The **analysis** directory contains:

-   [:file_folder: paper](/analysis/paper): Quarto source document for
    manuscript. Includes code to reproduce the figures and tables
    generated by the analysis.
-   [:file_folder: data](/analysis/data): Data used in the analysis.
-   [:file_folder: figures](/analysis/figures): Plots and other
    illustrations

## How to download and run locally in a Docker container

This research compendium includes a [Dockerfile](Dockerfile) that specifies the computational environment used to generate the results presented in the paper. You can recreate this environment, which includes R, RStudio and all necessary R packages, on your computer by following these steps:

1. Install [Docker](https://www.docker.com/get-started/) on your computer 
2. Download the compendium as a zip from from this URL:
[master.zip](/archive/master.zip) and unzip    
3. Set your terminal working directory to the compendium, and run `docker build -t wos .` to build the container on your computer, it will take a few minutes and require an internet connection.
4. Run in your terminal `docker run --rm -it -e ROOT=TRUE -e PASSWORD=rstudio -dp 8787:8787 wos` to run the container
5. Go to http://localhost:8787/ with your browser and enter rstudio in both fields to start RStudio in the browser, running in the Docker container
6. In RStudio, run `rstudioapi::openProject("/home/rstudio/web-of-science-archaeology")` to open this project
7. Run `rmarkdown::render("analysis/paper/paper.qmd")` to render the Quarto document
8. When finished, to remove the Docker container and files, run this in the terminal `docker ps -aq | xargs docker stop | xargs docker rm`

### Licenses

**Text and figures :**
[CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)

**Code :** See the [LICENSE](LICENSE.md) file

**Data :** [CC-0](http://creativecommons.org/publicdomain/zero/1.0/)
attribution requested in reuse

### Contributions

We welcome contributions from everyone. Before you get started, please
see our [contributor guidelines](CONTRIBUTING.md). Please note that this
project is released with a [Contributor Code of Conduct](CONDUCT.md). By
participating in this project you agree to abide by its terms.